### 2009-12-09 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
This revision incorporates the revisions from Niki for chapter 3, and my own corrections to chapters 1 and 2 using MS word to find misspelled words. Not it is a clean version.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@687 559edde2-a4d4-44e0-96fc-4035e40d7396

The property of coverage on a sensor network is defined by
\cite{sn-intro01} as the physical space ``visible" to the sensors devices, which
is also directly related to the purpose of the network, its size in terms of
the number of nodes deployed and the sensor network lifetime. Therefore,
sensors are deployed in sparse or dense manners, according to the
specification of what needs to be covered by the sensor devices. Furthermore,
considering that some types of sensor devices are very likely to experience regular
hardware failures, \cite{sn-intro02} described sensor network examples with 
redundant nodes that provide better data coverage.
Finally, the coverage of sensor devices also have influence on the connectivity
property of the sensor nodes. In order to communicate with redundant nodes, a
physical connection must exist at all times. On the other hand, lower
level of coverage is related to sporadic communication among the sensor nodes.

@@

The most basic computer application to process data is through the use of
spreadsheets, where data is organized into cells \cite{tabular-model}. When
exported to the so-called comma-separated values (CSV) file format, the
resulting file is comprised of lines of ASCII characters, separated by a
delimiter. The first line represents a header, with the identification of
columns of specific data, where the following lines contains the values
related to each top column. \cite{sn-provenance} reports that the tabular
data model is commonly used to provide data persistence for sensor networks.
The SF-BEAMS environmental sensor networks \cite{sfbeams2006} uses this
approach to distribute the collected data over the Internet using the OPenDAP
\cite{opendap} data format.

@@

Other examples of the use of the structured data model using XML is found in
the use of middlewares, as the authors of \cite{sn-xml-query-engines} reports.
Such an approach is used for data exchange by the implementing of a middleware
capable of query and extract the collected data from sensor devices
\cite{sn-xml-middleware}. A practical example of the use of XML as a middleware
to communicate between nodes is the communication mechanism used by NetBEAMS
\cite{netbeams2009}. See section \ref{sec:dsp-serialization} for details.

@@

In general, sensor devices produce and preserve data onto their local
storage or transmit them to data nodes as described earlier in this chapter. It
is vital to understand that the majority of these devices outputs the raw data
in streams of values with any type of delimitation between each other. For
this reason, the processes of parsing and labeling each of the values located
in the data stream regularly requires the use of the device's specification
documentation. Similarly, in order to have access to each of the properties of
collected data by a given property related to time or space,
\cite{sn-provenance} also mentions that Metadata\footnote{Data about data}
should be included together with the collected data.

@@

A different way to describe sensor devices is through the use of
Annotations. \cite{sn-annotation} shows the use of tags to describe data
collected from a physical area monitoring. The sensor network was
designed to monitor moving objects, capturing their location while in
movement and their visual appearance using a video camera. In this way,
different instances can be described by adding tags to particular video frames, and
therefore, providing an efficient way to search for a specific frame based on
its visual observations. Similarly, the above referenced project also uses data
provenance to describe its collected data, since each of the collected data 
contains the related GPS coordinates of the observed object.

@@

Different sensor devices produce different sets of attributes based on their
capabilities and design. In general, these data sets do not have distinct
names, which leads to difficulties when indices over the properties of the
collected data are required to make them searchable. Specially for the purpose
of data archival, the collected data needs to be identified, giving a meaning
of independence on its own, the same idea behind the unique identification of
entities on the relational data model \cite{relational-model}. Equally,
\cite{sn-provenance} describes that traditional sensor networks use the
collected data for analysis long after the attributes have been collected,
usually before it is categorized and indexed. Annotations and tags are also
names given by \cite{sn-provenance} as means to describe collected sensor data.

Briefly, the collected data is described by the name of the observation
property. For instance, \textbf{temperature} = 73.45 expresses the value of the
observation of the temperature. In other words, the string ``temperature'' is
the property describing the raw collected data 73.45. On the other hand,
the example \textbf{temperature-scale} = ``Fahrenheit": describes the metadata
regarding the temperature.

@@

\subsection{From where was the data collected?}

@@

Data management for large scales of data require different techniques to store
and retrieve data, since it requires more computational power and data
retrieval efficient algorithms. \cite{sn-data-center} describes the creation of
a Sensor Data Center designed to store and analyze amounts of data on the range of
Terabytes to a future Petabytes of data. Its sensor network deals with not only
numerical sensor readings from transportation management, but also radar
images, used for historic purposes and data mining that performs traffic
pattern analysis. In this way, the authors report the use of parallel data
processing techniques in order to provide processing for the collected
data, and load-balancing when storing data to minimize the sensor data
processing delays. For this reason, they suggest the use of concurrent control
such as snapshot isolation of the current data into separated replicated data.
Therefore, data access from users are not interrupted by the processing of
the data that is processed.

>>>>

The property of coverage on a sensor network is defined by
\cite{sn-intro01} as the physical space ``visible" to the sensors devices, which
is also directly related to the purpose of the network, its size in terms of
the number of nodes deployed and the sensor network lifetime. Therefore,
sensors are deployed in sparse or dense manners, according to the
specification of what needs to be covered by the sensor devices. Furthermore,
considering that some types of sensor devices are very likely to experience regular
hardware failures, \cite{sn-intro02} described sensor network examples with 
redundant nodes that provide better data coverage.
Finally, the coverage of sensor devices also has influence on the connectivity
property of the sensor nodes. In order to communicate with redundant nodes, a
physical connection must exist at all times. On the other hand, lower
level of coverage is related to sporadic communication among the sensor nodes.

@@

The most basic computer application to process data is through the use of
spreadsheets, where data is organized into cells \cite{tabular-model}. When
exported to the so-called comma-separated values (CSV) file format, the
resulting file is comprised of lines of ASCII characters, separated by a
delimiter. The first line represents a header, with the identification of
columns of specific data, where the following lines contain the values
related to each top column. \cite{sn-provenance} reports that the tabular
data model is commonly used to provide data persistence for sensor networks.
The SF-BEAMS environmental sensor networks \cite{sfbeams2006} uses this
approach to distribute the collected data over the Internet using the OPenDAP
\cite{opendap} data format.

@@

Other examples of the use of the structured data model using XML is found in
the use of middleware, as the authors of \cite{sn-xml-query-engines} reports.
Such an approach is used for data exchange by the implementing of a middleware
capable of query and extract the collected data from sensor devices
\cite{sn-xml-middleware}. A practical example of the use of XML as a middleware
to communicate between nodes is the communication mechanism used by NetBEAMS
\cite{netbeams2009}. See section the documentation at
\cite{netbeams-dsp-architecture} for details.

@@

In general, sensor devices produce and preserve data onto their local
storage or transmit them to data nodes as described earlier in this chapter. It
is vital to understand that the majority of these devices output the raw data
in streams of values with any type of delimitation between each other. For
this reason, the processes of parsing and labeling each of the values located
in the data stream regularly requires the use of the device's specification
documentation. Similarly, in order to have access to each of the properties of
collected data by a given property related to time or space,
\cite{sn-provenance} also mentions that Metadata\footnote{Data about data}
should be included together with the collected data.

@@

A different way to describe sensor devices is through the use of
Annotations. \cite{sn-annotation} shows the use of tags to describe data
collected from a physical area monitoring. The sensor network was
designed to monitor moving objects, capturing their location while in
movement and their visual appearance using a video camera. In this way,
adding tags to particular video frames can describe different them, and
therefore, providing an efficient way to search for a specific frame based on
its visual observations. Similarly, the above referenced project also uses data
provenance to describe its collected data, since each of the collected data 
contains the related GPS coordinates of the observed object.

@@

Different sensor devices produce different sets of attributes based on their
capabilities and design. In general, these data sets do not have distinct
names, which lead to difficulties when indices over the properties of the
collected data are required to make them searchable. Specially for the purpose
of data archival, the collected data needs to be identified, giving a meaning
of independence on its own, the same idea behind the unique identification of
entities on the relational data model \cite{relational-model}. Equally,
\cite{sn-provenance} describes that traditional sensor networks use the
collected data for analysis long after the attributes have been collected,
usually before it is categorized and indexed. Annotations and tags are also
names given by \cite{sn-provenance} as means to describe collected sensor data.

Briefly, the collected data is described by the name of the observation
property. For instance, \textbf{temperature} = 73.45 expresses the value of the
observation of the temperature. In other words, the string ``temperature'' is
the property describing the raw collected data 73.45. On the other hand,
the example \textbf{temperature-scale} = ``Fahrenheit" describes the metadata
regarding the temperature.

@@

\subsection{Data Origin}

@@

Data management for large scales of data require different techniques to store
and retrieve data, since it requires more computational power and data
retrieval efficient algorithms. \cite{sn-data-center} describes the creation of
a Sensor Data Center designed to store and analyze amounts of data on the range of
Terabytes to a future Petabytes of data. Its sensor network deals with not only
numerical sensor readings from transportation management, but also radar
images, used for historic purposes and data mining that performs traffic
pattern analysis. In this way, the authors report the use of parallel data
processing techniques in order to provide processing for the collected
data, and load balancing when storing data to minimize the sensor data
processing delays. For this reason, they suggest the use of concurrent control
such as snapshot isolation of the current data into separated replicated data.
Therefore, data access from users is not interrupted by the processing of
the data that is processed.

### 2009-12-09 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
This revision incorporates the revisions from Niki for chapter 3, and my own corrections to chapters 1 and 2 using MS word to find misspelled words. Not it is a clean version.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@687 559edde2-a4d4-44e0-96fc-4035e40d7396

As revailed in the previous chapter, data persistence for sensor
networks require an understanding of not only the properties of the
infrastructure of a given sensor network, but also of the properties and nature
of the collected data themselves. To put things differently, in order to
identify these concepts more clearly, each section of this chapter proposes an
individual taxonomy based on the aspects described in the literature
review. Then, it briefly discusses the relationships among them.

A taxonomy is defined as the practice and science of classification and uses
the taxonomic units of ``taxa'', the plural form of ``taxon''. Similarly, they
are represented by a hierarchy diagram that relates the main taxonomy to the
different taxa and their own related ones, if any.

@@

As delineated in section \ref{sec:sn-data-purpose}, the purpose of the
collected data from sensor devices determines how the collected data is
retrieved and used in the network. In such a fashion, this taxonomy relates to
how the collected data is used, and should be the primary question to be
answered during the analysis of a sensor network. The taxa related to this
taxonomy, depicted in Figure \ref{fig:taxonomy-data-purpose}, is described as
follows:

@@

The location where the collected data is stored plays an important role
in the different aspects of the access to the collected data from sensor
devices, as shown in section \ref{sec:sn-storage-locations}. In this way,
Figure \ref{fig:taxonomy-data-location} depicts this taxonomy, which can be
described as follows:

@@

\begin{itemize}
  \item \textbf{Local Storage}: characterizes the placed of the collected data
  temporarily located in-memory, or in a secondary storage device on a node
  located in-network;
  \item \textbf{External Storage}: The collected data is usually located in an
  external storage device with a data management system such as a relational
  database. When the collected data occupies different external storages, the
  strategy is called \textbf{Data-Centric Storage}, being organized by
  categories based on the collected data keys and values. In other words, what
  characterizes this taxon is the presence of a data partitioning strategy to
  store the collected data.
\end{itemize}

@@

One of the fundamental challenges in the development of a persistence storage
in sensor networks is related to the data model chosen to represent the
collected data. Different data models were reported in the review of related
literature in section \ref{sec:data-models}. As a consequence, this taxonomy
relates to the data model used to design the collected data, as drawn in Figure
\ref{fig:taxonomy-data-model}, and described as follows.

@@

\begin{itemize}
  \item \textbf{Schema-Dependent Models}: this data model requires the
  definition of a master schema that describes the data through a rigorous of 
  data modeling process prior to the insertion of the data. Examples of such
  model is the traditional Relational Data Model\cite{relational-model}, as
 well as the Structured Data Models such as the XML \cite{xml};
  \item \textbf{Schema-less Models}: contrary to the former taxon, schema-less
  data models do not require the definition of a data schema or table
  definition. Instances of such data model reviewed in the literature review
  is the use of the Tabular Data model.
\end{itemize}

@@

Particulaly related to how the collected data is described, different projects
provide data persistence describing the collected data by using its natural
properties, or associating textual descriptions or keywords of the data. The
way to describe data can follow different guidelines such as data provenance
and annotations, discussed in section \ref{sec:sn-data-description}. As
reported in section \ref{sec:sn-data-description}, the description of the
properties should take into account different properties of the nature of the
data and any other description needed. Therefore, the taxa that composes the
Data Description taxonomy, shown in Figure
\ref{fig:taxonomy-data-description}, are described as follows:

@@

\begin{itemize}
  \item \textbf{Provenance-Aware}: when the collected data is categorized based
  on the nature of the data, \textbf{What: Data Identity} is the data type
  that uniquely identifies the collected data from sensors. Similarly, the
  \textbf{When: Time Dimension}, is related instants of time related to the
  collected data: \textbf{fact} and \textbf{transaction} times. The former is
  used to identify when the data was collected, whereas the latter when the data was
  transmitted to the data sink where the query processing acts. Finally, the
  \textbf{Where: Data Location} identifies the point where the collected data
  was observed by the sensor device. Usually the latitude and longitude as the
  GPS coordinates;
  \item \textbf{Annotations}: the annotation of the collected data, through the
  use of keywords, full-text description, or any other method that helps
  describing the collected data. It is important to note that this type of data
  does not make part of the collected data from sensors. Rather, they
  complement the observations of what was percieved by the sensor devices.
\end{itemize}

@@

The query processing mechanism in sensor networks is directly related to how
and where the collected data is used and stored, as shown in section
\ref{sec:query-process}. As a consequence, the following taxa are related to the
different types of query processing mechanisms and summarized in Figure
\ref{fig:taxonomy-query-mechanism}:

@@

\begin{itemize}
  \item \textbf{In-Network Query Processing}: when the data node is the
  sensor device itself, and it uses the local storage mechanism, then the
  sensor node is able to reply to request related to the observations of the 
  environment defined by the sensor device. In this way, the query takes 
  place in the network itself, and usually returns values related to the current
  state of the environment;
  \item \textbf{Centralized Query Processing}: as developed by a single 
  relational database server, the data queried from a centralized data
  network sink in order to be reused. In general, this approach is used when
  the purpose of data is archival data.
\end{itemize}

@@

When it comes to the data volume produced during a time frame, different ranges
were reported from different types of sensor networks. Depending on the types
of data produced such as numerical or images, they can be in the orders of
Megabytes (1MB = 1000KB = ~10\superscript{3} bytes), Gigabytes (1GB =
1000\superscript{3} = 10\superscript{9} bytes) and even Terabytes (1TB =
1000\superscript{4} = 10\superscript{12}) to future Petabytes (1PB =
1000\superscript{5} = 10\superscript{15} bytes) as shown in section
\ref{sec:data-load}. As a comparison of what is processed, Google reports to
process 20 Petabytes of data per day in their data centers
\cite{map-reduce-load}. In this way, the following taxa are related to the
different sizes of data produced by a sensor network, summarized in Figure
\ref{fig:taxonomy-data-volume}:

@@

\begin{itemize}
  \item \textbf{Small}: sensor networks whose yearly produced data is in the
  range of MBs or to 1GB of data, usually stored in flat files directly in the
  operating system. They are usually collected from sensor devices that only
  generate numerical data;
  \item \textbf{Medium}: amount of data that ranges from GBs to 1TB data. In
  order to manage such amount of data using disk arrays;
  \item \textbf{Large}: when the amount of data produced by a sensor device
  reaches amounts of data compatible to data centers, in the orders ranging
  from TBs to 1PT of data. The construction of a data center specialized for
  data produced by sensor devices, taking into account the addition of data
  provenance, among others.
\end{itemize}

@@

Different approaches of are developed computational systems to deal the data
produced from sensor networks. As  
In this way, Figure \ref{fig:taxonomy-database-architecture} describes the
taxonomy of the database system organization as follows:

@@

\begin{itemize}
  \item \textbf{Centralized System}: a database system that runs in one single
  centralized host, being the focal purpose managing the collected data
  \cite{sn-intro01};
  \item \textbf{Distributed System}: a database system that can be composed by
  more than one node. Usually, the use of the pattern of Master-Slave
  mechanism or data ``Replication" is used to aliviate ``hot spots'' during
  concurrent processes of data collection from the sensor devices and data
  access from the final users. On the other hand, unique mechanisms can be
  used to disagragate the produced data from a concentrated one: the use 
  of database partioning, as related to the data-centric storage, where  
  each partition is responsible to hand data produced by a given type of 
  sensor, for instance.
\end{itemize}

@@

On the whole, the taxonomies described in this chapter were based solely on the
literature review of described in the previous chapter. They can be helpful in 
understanding of requirements of a data persistence layer for sensor networks
with different infrastructures. They are summarized on table
\ref{tab:taxonomies-list-summary}, and briefly related highlight these
relationships in order to better analyze a problem related to data persistence.

@@

The taxonomy related to the purpose of the sensor data can be directly related
to the taxonomy of the location of the sensor data. While the taxon of the 
real-time data stream relates to the local data storage, the taxon related to
data archival correlates the storage of the collected data with the an external
storage one. In addition to that, the converse is also true.

The taxonomy related to the data model can be related to any taxon from the
taxon described from the system organization, to data description and query
processing and location of the collected data. In the former, the data of a
relational model is stored using SQL messages.

Similarly, the taxonomy related to the volume of data produced by a sensor
network usually determines the system organization used to handle and process
the collected data. When the volume is small, the use a single server is
required, usually using an external storage device, which relates to the
taxonomy for storage system. However, when the volume is medium to large, the
use of distributed systems is required to cope with high volume of data
processing. Techniques described for data replication is used to decrease the
load in hot spots of the architecture. On the other hand, approaches to
separate the data being written and read can be used such as Load Balancing or
Database Partition. As a consequence, the use of a Data-Centric storage
approach is used.

In the light of these taxonomies, the classification of the data persistence
properties of a sensor networks can be studied and reviewed. As described in
the introduction of this dissertation, one the main motivations for this work
was to provide means of data persistence for NetBEAMS, a case study which does
not contain a data layer. Therefore, the following chapter seeks to describe
NetBEAMS and its supporting infrastructure.

>>>>

As reveled in the previous chapter, data persistence for sensor
networks require an understanding of not only the properties of the
infrastructure of a given sensor network, but also of the properties and nature
of the collected data themselves. In order to
identify these concepts more clearly, each section of this chapter proposes an
individual taxonomy based on the aspects described in the literature
review. Then, it briefly discusses the relationships among them.

Taxonomy is defined as the practice and science of classification and uses
the taxonomic units of ``taxa'', the plural form of ``taxon'' \cite{taxonomy}.
Similarly, they are represented by a hierarchical diagram that relates the
main taxonomy to the different taxa and their own related ones, if any.

@@

As delineated in section \ref{sec:sn-data-purpose}, the purpose of the
collected data from sensor devices determines how the collected data is
retrieved and used in the network. In such a fashion, this taxonomy relates to
how the collected data is used, and should be the primary question answered 
during the analysis of a sensor network. The taxa related to this
taxonomy, depicted in Figure \ref{fig:taxonomy-data-purpose}, is described as
follows:

@@

The location where the collected data is stored plays an important role
in the different aspects of the access to the collected data from sensor
devices, as shown in section \ref{sec:sn-storage-locations}. Figure 
\ref{fig:taxonomy-data-location} depicts this taxonomy, which can be
described as follows:

@@

\begin{itemize}
  \item \textbf{Local Storage}: characterizes the place the collected data is
  temporarily located in-memory, or in a secondary storage device on a node
  located in-network;
  \item \textbf{External Storage}: The collected data is usually located in an
  external storage device with a data management system such as a relational
  database. When the collected data occupies different external storage devices, the
  strategy is called \textbf{Data-Centric Storage}, being organized by
  categories based on the collected data keys and values. In other words, what
  characterizes this taxon is the presence of a data partitioning strategy to
  store the collected data.
\end{itemize}

@@

One of the fundamental challenges in the development of a persistence storage
in sensor networks is related to the data model chosen to represent the
collected data. This taxonomy relates to the data model used to design the 
collected data, as drawn in Figure \ref{fig:taxonomy-data-model}, and is 
described as follows.

@@

\begin{itemize}
  \item \textbf{Schema-Dependent Models}: this data model requires the
  definition of a master schema that describes the data through a rigorous
  data modeling process prior to the insertion of the data. Examples of this
  model is the traditional Relational Data Model\cite{relational-model}, as
 well as the Structured Data Models such as the XML \cite{xml};
  \item \textbf{Schema-less Models}: contrary to the former taxon, schema-less
  data models do not require the definition of a data schema or table
  definition. Instances of such a data model is the use of the 
  Tabular Data model.
\end{itemize}

@@

Particularly related to how the collected data is described, different projects
provide data persistence describing the collected data by using its natural
properties, or associating textual descriptions or keywords for the data. The
way to describe or label data can follow different guidelines such as data provenance
and annotations, discussed in section \ref{sec:sn-data-description}. The description of the
properties should take into account the different values of the nature of the
data and any other description needed. Therefore, the taxa that composes the
Data Description taxonomy, shown in Figure
\ref{fig:taxonomy-data-description}, are described as follows:

@@

\begin{itemize}
  \item \textbf{Provenance-Aware}: when the collected data is categorized based
  on the nature of the data, \textbf{What: Data Identity} is the data type
  that uniquely identifies the collected data from sensors. Similarly, the
  \textbf{When: Time Dimension}, is related to instances of time relative to the
  collected data: \textbf{fact} and \textbf{transaction} times. The former is
  used to identify when the data was collected, whereas the latter when the data was
  transmitted to the data sink where the query processing acts. Finally, the
  \textbf{Where: Data Location} identifies the point where the collected data
  was observed by the sensor device. Usually the latitude and longitude as the
  GPS coordinates;
  \item \textbf{Annotations}: the annotation of the collected data, through the
  use of keywords, full-text description, or any other method that helps
  describe the collected data. It is important to note that this type of data
  does not makeup part of the collected data from sensors. Rather, they
  complement the observations of what was perceived by the sensor devices.
\end{itemize}

@@

The query processing mechanism in sensor networks is directly related to how
and where the collected data is used and stored, as shown in section
\ref{sec:query-process}. The following taxa are related to the
different types of query processing mechanisms and summarized in Figure
\ref{fig:taxonomy-query-mechanism}:

@@

\begin{itemize}
  \item \textbf{In-Network Query Processing}: when the data node is the
  sensor device itself, and it uses the local storage mechanism, then the
  sensor node is able to reply to the request related to the observations of the 
  environment defined by the sensor device. In this way, the query takes 
  place in the network itself, and usually returns values related to the current
  state of the environment;
  \item \textbf{Centralized Query Processing}: as developed by a single 
  relational database server, the data queried from a centralized data
  network sink in order to be reused. In general, this approach is used when
  the purpose of data is archival data.
\end{itemize}

@@

When it comes to the data volume produced during a time frame, different ranges
are reported from different types of sensor networks. Depending on the types
of data produced such as numerical or images, they can be in the orders of
Megabytes (1MB = 1000KB = ~10\superscript{3} bytes), Gigabytes (1GB =
1000\superscript{3} = 10\superscript{9} bytes) and even Terabytes (1TB =
1000\superscript{4} = 10\superscript{12}) to future Petabytes (1PB =
1000\superscript{5} = 10\superscript{15} bytes) as shown in section
\ref{sec:data-load}. As a comparison of what is processed, Google reports to
process 20 Petabytes of data per day in their data centers
\cite{map-reduce-load}. In this way, the following taxa are related to the
different sizes of data produced by a sensor network, summarized in Figure
\ref{fig:taxonomy-data-volume}:

@@

\begin{itemize}
  \item \textbf{Small}: sensor networks whose annual production of data is in the
  range of MBs or to 1GB of data, usually stored in flat files directly in the
  operating system. They are usually collected from sensor devices that only
  generate numerical data;
  \item \textbf{Medium}: amount of data that ranges from GBs to 1TB data and is 
  managed using disk arrays;
  \item \textbf{Large}: the amount of data produced by a sensor device ranging 
  from TBs to 1PT of data necessitates the construction of a data center 
  specializing in the storage of data produced by sensor devices, taking into 
  account the addition of data provenance, among others.
\end{itemize}

@@

Different approaches have been developed to organize the storage of data
produced from sensor networks. Figure \ref{fig:taxonomy-database-architecture} 
describes the taxonomy of the database system organization as follows:

@@

\begin{itemize}
  \item \textbf{Centralized System}: a database system that runs in one single
  centralized host, the focus being the managing of the collected data
  \cite{sn-intro01};
  \item \textbf{Distributed System}: a database system that can be composed of
  more than one node. Usually, the use of the pattern of Master-Slave
  mechanism or data ``Replication" is used to alleviate ``hot spots'' during
  concurrent processes of data collection from the sensor devices and data
  access from the final users. Unique mechanisms can be used to disaggregate 
  the produced data from a concentrated one: the use of database partitioning, 
  as related to the data-centric storage, where each partition is responsible 
  to handle data produced by a given type of sensor, for instance.
\end{itemize}

@@

On the whole, the taxonomies described in this chapter were based solely on the
literature review described in the previous chapter. They can be helpful in 
understanding the requirements of a data persistence layer for sensor networks
with different infrastructures. They are summarized on Table
\ref{tab:taxonomies-list-summary}, and briefly highlight these
relationships in order to better analyze a problem related to data persistence.

@@

The taxonomy related to the purpose of the sensor data can be directly associated 
with the taxonomy of the location of the sensor data. While the taxon of the 
real-time data stream relates to the local data storage, the taxon related to
data archival correlates the storage of the collected data with an external
storage one. In addition to that, the converse is also true.

The taxonomy related to the data model can be linked to any taxon from the
taxon described from the system organization, to data description and query
processing and location of the collected data. In the former, the data of a
relational model is stored using SQL messages.

Similarly, the taxonomy related to the volume of data produced by a sensor
network usually determines the system organization used to handle and process
the collected data. When the volume is small, the use a single server is
required, usually using an external storage device, which relates to the
taxonomy for storage system. However, when the volume is medium to large, the
use of distributed systems is required to cope with high volume of data
processing. Techniques described for data replication is used to decrease the
load in hot spots of the architecture. On the other hand, approaches to
separate the data being written and read can be used such as Load Balancing or
Database Partition, and the use of a Data-Centric storage approach is used.

In the light of these taxonomies, the classification of the data persistence
properties of a sensor networks can be studied and reviewed. As described in
the introduction of this report, one the main motivations for this work
was to provide means of data persistence for NetBEAMS, a case study which does
not contain a data layer. Therefore, the following chapter seeks to describe
NetBEAMS and its supporting infrastructure.
